---
title: 'Reproducible Research: Peer Assessment 1'
author: "Christophe Guerrier"
date: "17 October 2014"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
First we unzip the data, and load it with read.csv.
```{r, echo=TRUE }
unzip(zipfile = "./activity.zip")
activity = read.csv(file = "./activity.csv",header=TRUE)
```
Few useful package
```{r, echo=TRUE }
library(dplyr)
library(ggplot2)
```
Now converting date, and compute time from intervas.
```{r, echo=TRUE }
activity$day = as.Date(activity$date,format = "%Y-%m-%d")
activity$hour = floor(activity$interval/100)
activity$min = 100 *(activity$interval /100 - floor(activity$interval/100))
activity$time = as.POSIXct(x = paste(activity$date,activity$hour,activity$min),format = "%Y-%m-%d %H %M")
```
Compute the number of steps for each day.
```{r, echo=TRUE }
stepsDay <- colSums(table(activity$steps, activity$date))
```


## What is mean total number of steps taken per day?
Histogram of the total number of steps taken each day
```{r, echo=TRUE }
hist(stepsDay, breaks=50)
```

```{r, echo=TRUE }
mean(stepsDay)
median(stepsDay)
```

## What is the average daily activity pattern?
```{r, echo=TRUE }
plot(x = activity$time, y=activity$steps, type = 'l',
     main = "Time series, steps activity", ylab = "Steps", xlab="Time")
```
Looking at daily activity.






## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?